Safe and Robust Formal Synthesis of Digital Controllers for Continuous Plants

### 2018-02-27 automatica/paper.tex
sections 2 and 3

\textbf{Discretization Effects -}
The classical approach to control synthesis has often disregarded
digitalization effects, whereas more recently modern techniques have focused
on different aspects of discretization, including delayed
response~\cite{Duggirala2015} and finite word length (FWL) semantics, with
the goal either to verify (e.g.,~\cite{daes20161}) or to optimize
(e.g.,~\cite{oudjida2014design}) given implementations.

@@

We consider models of physical plants expressed as ordinary differential
equations (ODEs), which we assume to be controllable~\cite{Astrom08,
astrom1997computer}: \note{Pls eliminate vector quantities, unless strictly needed. What do you mean by vector quantities?}
%
\begin{align}
\label{eq:ode1}
\dot{x}(t) = \mat{A}\vec{x}(t)+ \mat{B} \vec{u}(t), \quad 
\vec{x} \in \mathbb{R}^p, \vec{u} \in \mathbb{R}^q,\,
\mat{A} \in \mathbb{R}^{p \times p},\,
\mat{B} \in \mathbb{R}^{p \times q},
\end{align}
%
where $t \in \mathbb R_0^+$ denotes continuous time, $\mat{A}$ and $\mat{B}$
are matrices that fully characterize the dynamics of the continuous plant,
and $x(0)$ is the initial state.

@@

%-------------------------------
\subsection{Abstract Acceleration} 
\label{ssec:aa}
%-------------------------------

For abstraction based synthesis, we will use Abstract
Acceleration~\cite{JSS14,cattaruzza2015unbounded}.  Given an iterative
program with dynamics
%
\begin{equation}
\vec{x}_{k+1}=\mat{A}\vec{x}_k : \vec{x}_k \in \mathbb{R}^n \wedge \mat{A} \in \mathbb{R}^{n \times n} \wedge k \in [0\ \ \infty),
\end{equation}
%
we want to find all possible states visited over an infinite time horizon
starting at an initial set $X_0$.  We call this set the \emph{reach tube} of
the model:
%
\begin{equation}
\hat{X} = \{ \vec{x}_k : k \in [0\ \infty) \wedge \vec{x}_0 \in X_0 \wedge \vec{x}_{k+1}=\mat{A}\vec{x}_k\}.
\end{equation}
%
It is easy to see that the infinite number of iterations would result in an
unbounded cumulative rounding error, but using Abstract Acceleration we
transform this problem into a one-step solution.  We first change the
problem statement by means of acceleration into
%
\begin{equation}
\hat{X} = \{ \vec{x}_k=\mat{A}^k\vec{x}_0 : k \in [0\ \infty) \wedge \vec{x}_0 \in X_0 \}, 
\end{equation}
%
and further define new model semantics and corresponding set: 
%
\begin{equation}\label{eq:aa_reachtube}
\hat{X} \subseteq \hat{X}^\sharp = \mathcal{A}X_0, \text{ such that } \bigcup_{k \in [0\ \infty)} \mat{A}^k \subseteq \mathcal{A}, 
\end{equation}
%
where the newly introduced \emph{abstract reach tube} $\hat{X}^\sharp$ is an
over-approximation of the reach tube of the model \cite{JSS14}.  All we need
to do is define the nature of $\mathcal{A}X_0$.

@@

Introduce set $X=\{ \vec{x} \mid \mat{C}\vec{x} \leq \vec{d}\}$. %\note{This notation clashes with that of the output map for the LTI model. }
The linear transformation $\vec{x}'=\mat{S}^{-1}\vec{x}$ leads to set 
%
\begin{equation}
X'= \mat{S}^{-1}X = \{\vec{x}' : \mat{C}\mat{S}\vec{x}'\leq \vec{d}\}.  
\label{eq:eigenInit}
\end{equation}
%
Analogously, let $X_0'=\mat{S}^{-1}X_0$ and
${\hat{X}'^\sharp}=\mat{S}^{-1}{\hat{X}^\sharp}$ be the initial set and the
abstract reach tube mapped onto the eigenspace of matrix $\mat{A}$.  We may
transform equation \eqref{eq:aa_reachtube} into: 
%
\begin{equation}
\hat{X}' \subseteq {\hat{X}'^\sharp} = \mathcal{J}X_0', 
\end{equation}
%
where $\hat{X}'=\mat{S}^{-1}\hat{X}$ and $\mathcal{J}$ is a convex
polyhedron representing restrictions on the eigenvalues of $\mat{A}$. 
$\hat{X}'^\sharp$ is calculated by applying a simplex algorithm with a
tableau defined by $\mathcal{J}$ and objective functions derived from each
vertex of $X_0'$ (step \eqref{alg:getReachTube} in Algorithm
\ref{alg:findReachTube} below).  We note, however, that if we wish to use
numerical analysis in the decomposition of $\mat{A}$, the Tableau in
$\mathcal{J}$ contains small intervals given by the errors, which cannot be
processed efficiently by a regular simplex and require a new procedure.

>>>>

\textbf{Discretization Effects -}
The classical approach to control synthesis has often disregarded
digitalization effects, whereas more recently modern techniques have focused
on different aspects of discretization, including delayed
response~\cite{Duggirala2015} and finite word length (FWL) semantics, with
the goal either to verify~\cite{daes20161} or to optimize~\cite{oudjida2014design} given implementations.

@@

We consider models of physical plants expressed as ordinary differential
equations, which we assume to be controllable~\cite{Astrom08,
astrom1997computer}: \note{Pls eliminate bold quantities, unless strictly needed. }
%
\begin{align}
\label{eq:ode1}
\dot{x}(t) = \mat{A}\vec{x}(t)+ \mat{B} \vec{u}(t), \quad 
\vec{x} \in \mathbb{R}^p, \vec{u} \in \mathbb{R}^q,\,
\mat{A} \in \mathbb{R}^{p \times p},\,
\mat{B} \in \mathbb{R}^{p \times q},
\end{align}
%
where $t \in \mathbb R_0^+$ denotes continuous time, $\mat{A}$ and $\mat{B}$
are matrices that fully characterize the dynamics of the continuous plant,
and $x(0)$ is the initial state.

@@

%-------------------------------
\subsection{Abstract Acceleration of Linear Dynamics} 
\label{ssec:aa}
%-------------------------------

We recall a technique called \emph{Abstract Acceleration}~\cite{JSS14,cattaruzza2015unbounded}, 
which will be key for abstraction-based synthesis.  
\note{This clashes with the model introduced above, with inputs - motivate why?}
Given a model of an iterative program with dynamics
\note{Pls eliminate unnecessary ``logical'' notation - we have discussed this already for the CAV18 submission, do we need to go through this again????}
%
\begin{equation}
\vec{x}_{k+1}=\mat{A}\vec{x}_k : \vec{x}_k \in \mathbb{R}^n \wedge \mat{A} \in \mathbb{R}^{n \times n} \wedge k \in \mathbb N_0,
\end{equation}
%
\note{material has been blindly pasted here - dimensions are all screwed up.}
we want to find all possible states visited over an infinite time horizon
starting at an initial set $X_0$.  
This set is known as the \emph{reach tube} of
the model:
%
\begin{equation}
\hat{X} = \{ \vec{x}_k : k \in \mathbb N_0 \wedge \vec{x}_0 \in X_0 \wedge \vec{x}_{k+1}=\mat{A}\vec{x}_k\}.
\end{equation}
%
It is easy to see that the infinite number of iterations would result in an
unbounded cumulative rounding error \note{why do we bring in the discussion on the rounding error here? why do we claim AA solves it? Is this the best manner to motivate the AA technique?}, but using Abstract Acceleration we
transform this problem into a one-step solution.  We first change the
problem statement by means of acceleration into
%
\begin{equation}
\hat{X} = \{ \vec{x}_k=\mat{A}^k\vec{x}_0 : k \in [0\ \infty) \wedge \vec{x}_0 \in X_0 \}, 
\end{equation}
%
and further define new model semantics and corresponding set: 
%
\begin{equation}\label{eq:aa_reachtube}
\hat{X} \subseteq \hat{X}^\sharp = \mathcal{A}X_0, \text{ such that } \bigcup_{k \in [0\ \infty)} \mat{A}^k \subseteq \mathcal{A}, 
\end{equation}
%
where the newly introduced \emph{abstract reach tube} $\hat{X}^\sharp$ is an
over-approximation of the reach tube of the model \cite{JSS14}.  All we need
to do is define the nature of $\mathcal{A}X_0$.

@@

Introduce set $X=\{ \vec{x} \mid \mat{C}\vec{x} \leq \vec{d}\}$. 
\note{Why are we interested in this set? What does this set represent?}
The linear transformation $\vec{x}'=\mat{S}^{-1}\vec{x}$ leads to set 
%
\begin{equation}
X'= \mat{S}^{-1}X = \{\vec{x}' : \mat{C}\mat{S}\vec{x}'\leq \vec{d}\}.  
\label{eq:eigenInit}
\end{equation}
%
Analogously, let $X_0'=\mat{S}^{-1}X_0$ and
${\hat{X}'^\sharp}=\mat{S}^{-1}{\hat{X}^\sharp}$ be the initial set and the
abstract reach tube mapped onto the eigenspace of matrix $\mat{A}$.  We may
transform equation \eqref{eq:aa_reachtube} into: 
%
\begin{equation}
\hat{X}' \subseteq {\hat{X}'^\sharp} = \mathcal{J}X_0', 
\end{equation}
%
where $\hat{X}'=\mat{S}^{-1}\hat{X}$ and $\mathcal{J}$ is a convex
polyhedron representing restrictions on the eigenvalues of $\mat{A}$. 
$\hat{X}'^\sharp$ is calculated by applying a simplex algorithm with a
tableau defined by $\mathcal{J}$ and objective functions derived from each
vertex of $X_0'$ (step \eqref{alg:getReachTube} in Algorithm
\ref{alg:findReachTube} below).  We note, however, that if we wish to use
numerical analysis in the decomposition of $\mat{A}$, the Tableau in
$\mathcal{J}$ contains small intervals given by the errors, which cannot be
processed efficiently by a regular simplex and require a new procedure.
